---
title: "Residualise m/beta-values"
author: X Shen  
date: "\n`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r library,echo=F}
library(dplyr)
library(readr)
```

```{r rendering docs, eval=F,echo=F}

```

------------------------------------------------------------------------

This protocol is created for (1) covarying and (2) creating methylation PCs using m/beta-values. You could choose to use either of the utilities or both.

The function uses multiple files or a single file.

Contact: xueyi.shen\@ed.ac.uk

------------------------------------------------------------------------

### Required R packages

-   dplyr

-   plyr

-   FactoMineR

-   optparse

### Function preparation

-   Download the R function from [here](https://github.com/psychiatric-genomics-consortium/mdd-mwas/util/residualise_mvalue.R)

------------------------------------------------------------------------

### File preparation

#### Methylation data

-   **rds** format

-   Row names: CpG probe names (typically start with 'cg'). Column names: Subject ID

-   Data should look like (italic cells are either column/row names):

|              | *Subject1* | *Subject2* | *Subject3* | *...* |
|:------------:|:----------:|:----------:|:----------:|:-----:|
| *cg08730728* |  -0.0012   |  -0.07376  |  0.47836   |  ...  |
| *cg03070533* |  0.13523   |  -0.81432  |  0.24235   |  ...  |
| *cg21588135* |  -0.7958   |  0.000257  |  -0.00532  |  ...  |
|    *...*     |    ...     |    ...     |    ...     |  ...  |

#### Covariates

-   **rds** format

-   First column should be participant IDs, consistent with methylation IDs.

-   Suggested covariates: age, sex, batch, assessment centre (please include as many as you can)

Data should look like (italic cells are either column/row names):

|   *ID*   | *age* | *sex*  | *batch* |
|:--------:|:-----:|:------:|:-------:|
| Subject1 |  45   |  Male  |    3    |
| Subject2 |  63   | Female |    1    |
| Subject3 |  60   | Female |    1    |
|   ...    |  ...  |  ...   |   ...   |

------------------------------------------------------------------------

### Run the script

#### Example scripts (running from bash command line)

For a single file:

```{bash, eval=F}
Rscript resid_pc_mvalue.R \
--meth data/mvalue.chr22.rds \
--covar covs.rds \
--residualise y \
--makePC y \
--out data/
```

For a list of files (large memory required; example file names: data/mvalues.chr1.rds, data/mvalues.chr2.rds, ...):

```{bash, eval=F}
Rscript resid_pc_mvalue.R \
--meth data/mvalue.chr \
--covar covs.rds \
--residualise y \
--makePC y \
--out data/
```

#### Process(es)

-   --residualise: y/n (y = residualise methylation data against covariates and save residualised data)

-   --makePC: y/n (y = make methylation PCs)

#### Output(s)

In the given output directory:

-   A list of files for residualised data (end with *\_resid.rds*), if 'resid' is provided for the 'process' option

-   A file called *methPCs.rds*, if 'pc' is provided for the 'process' option.
